The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2023-07-09 carbon_paper/tex/main.tex
thesis edits

\graphicspath{{../figures/}}

@@

\newcommand{\Yct}{\ensuremath{y_{\rm C}}}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm cc}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm cc}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm agb}}}

    \newcommand{\y}{\ensuremath{\rotatebox[origin=B,y=0.5ex]{180}{y}}}
\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else M$_{\sun}$
    \fi}
\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}

@@

After a single stellar population (SSP)\footnotemark{} forms, \cc{} are the first chemical enrichers. \cc{} explode within $\lesssim 40$\,Myr, providing light elements (C, O, and Mg) and heavier elements (Fe and beyond). Next, low-mass stars begin to reach the end of their lives. By shedding their outer layers, \agb{} stars are important sources of C, N, and neutron capture elements.  Finally, white dwarfs explode in \ia{}, releasing Fe and other iron-peak elements.

@@

To quantify yields, we define the stellar yield to be the fraction of a stars initial mass which is newly synthesized and released as a given element. 
For an element $X$ and star with mass $M$, the net-fractional stellar yield $\y$ is 
\begin{equation}
\y_{X} =  \Delta Z_X \frac{M_{\rm ejected}}{M_{\rm birth}}
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm birth}$  are the total ejected mass and the birth mass of the star, and $\Delta Z_X$ is the change in mass fraction of abundance X from birth to ejected. For example, if a 1\,\Mo\ star has $\y_{\C} = 10^{-3}$, then the star will add $10^{-3}\,\Mo$ of new C to the interstellar medium. 
Although per-star yields are necessary to compute \agb{} star enrichment rates in \gce{}  models, \imf-averaged yields are useful in interpreting their predictions. 
Given the fractional yield $\y(M, Z)$ as a function of initial stellar mass $M$ and metallicity $Z$, the \imf-averaged yield is given by 
\begin{equation} \label{eq:imf-yield}
    y_{\rm X}(Z,t) = 
    \frac{
    \int_{M_{\rm min}(t)}^{M_{\rm max}} 
    \y_{\rm X}(M, Z)
    \frac{dN}{dM}\ M\ dM
}
{
    \int \frac{dN}{dM}\ M\ dM
}
\end{equation}
where ${dN}/{dM}$ is the \imf, $M_{\rm max}=100\,\Mo$ is the maximum stellar mass, and $M_{\rm min}(t)$ is the mass of stars with lifetime $t$.\footnotemark{}
We use $t=10\,$Gyr for total yields when $t$ is not used.
To calculate the \imf-averaged net yields, we use the Versatile Integrator for Chemical Evolution code (\VICE\footnotemark{}).

@@

	\begin{tabular}{l l l l}
		\hline
        Element & $y^{\rm cc}$ & $\y^{\rm agb}$ & $y^{\rm ia}$ \\
		\hline
        C & Eq.~\ref{eq:zeta} & $2.9\times{\rm C11}$ &  0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than about??2\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
In 2?? to 4?? \Mo stars, third dredge up becomes important, enriching the outer layers with C, making these stars the most important producers of C. 
In massive, high-metallicity \agb\ stars, both hot bottom burning and third dredge up occur; however, third dredge up is much more efficient, resulting in significant \C[12] destruction.

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following.
\begin{enumerate}
    \item \cxi: \citet{cristallo+11, cristallo+15}
    \item \kx: \citet{karakas10}
    \item \vxiii: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \kxvi: \citet{KL16, karakas+18}
\end{enumerate}
In Appendix~\ref{sec:oob_models}, the masses, metallicites, and notable differences between each yield set are briefly described. 
For our models to match observations, we find that need to uniformly amplify these yield tables. We use \cxi\ table, amplified by a factor of 2.9, as the fiducial \agb\ yield.

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yield for these four models.
Note that the net yields may be negative if a star ejects material with a lower average C abundance than the material the star was formed from.
Most models agree on the qualitative shape of the net fractional \agb\ C yield---% 
stars with masses between about 2--4\,\Mo have the highest fractional C yields. As metallicity increases, the total \agb\ C yield decreases and the mass of the peak increases.

@@

The right panel of Fig.~\ref{fig:agb-ssp} shows \imf-averaged C yields for each \agb\ model as a function of metallicity.
\vxiii{} differs in that it shows a non-monotonic metallicity dependence. However, this effect is only for models with $\log Z/Z_\odot \lesssim -1$.
Otherwise, models differ only in their yield normalization and metallicity dependence. All models predict yields within a factor of \about{2} for fixed metallicity.
For example, the three models \cxi, \kx{}, and \kxvi{} predict $y_\text{C}^\text{agb}$ to be between 0.006 and 0.008 at solar metallicity, but \cxi\ has a much shallower metallicity dependence than the \kx{} and \kxvi{} models. \vxiii{} instead predicts a yield \about{0.004}.
    As shown in the right panel of Fig.~\ref{fig:agb-ssp}, with increasing $Z$, C enrichment occurs earlier, and C destruction in low-mass stars leads to a declining C abundance at late times.

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process. However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the many stellar modeling uncertainties.

Fig.~\ref{fig:y_cc} plots calculations of the \imf-integrated yields, defined with Eq.~\ref{eq:imf-yield} (computed using \VICE's \texttt{vice.yields.ccsne.fractional} function). 
\cc{} models predict a wide range of C yields, spanning almost a factor of ten. 
Both the \citet{NKT13} and \cite{LC18} models show positive metallicity dependence. 
The \cite{LC18} models also include rotation, showing that variations in the rotational velocity of the star can dramatically increase the magnitude and metallicity dependence of $\Ycc$. Rotation induces more mixing allowing the CO core to grow larger. As we will later show, \cc\ C production needs to be strongly metallicity-dependent at $Z/Z_\odot \approx 1$, which is consistent with the \cite{LC18} rapidly rotating models.

Fig.~\ref{fig:y_cc} shows the \cxi{} \agb\ model for comparison on the left. Especially at $Z\approx Z_\odot$, most \cc{} models dominate \agb\ C production. Later, we will also show empirically this is the case.
On the right of Fig.~\ref{fig:y_cc}, we also show the \cc{} [C/Mg] ratio, defined by
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/Mg]^{CC}} = \log_{10}\left( \frac{\Ycc}{\Yoc}\right) - \log_{10} \left( \frac{Z_{{\rm C},\ \sun }}{Z_{{\rm Mg},\ \sun }} \right).
\end{equation}
If only \cc{} produced C, then ${\rm [C/Mg]^{CC}}$ describes the equilibrium abundance of [C/Mg].
Once again, different \cc\ models span a large range in [C/Mg]. 
we chose to instead parameterize $\Ycc$ to enable agreement with observations, as most \cc\ models fail to achieve near-solar [C/Mg].
Assumptions about the explodability landscape affect C and Mg production. Increasing the fraction of stars that explode increases $\Ycc$, as stars that directly collapse do not contribute to explosive yields \citep{emily+21}. However, C is relatively unaffected by the black-hole landscape, as very massive stars contribute C through enriched winds. Since Mg is formed deeper in the core of \hms, the Mg yield drops much more steeply, so models where few stars explode (S16/W18) have higher [C/Mg].

\cc\ models do not reach [O/Mg] due to overproduction of O or underproduction of Mg, or both \citep{emily+21}. Here, we assume ${\rm [O/Mg]} = 0$, which is not compatible with \cc\ models but is consistent with \apogee{} observations \citep{weinberg+19, weinberg+22}.

Out model predictions of [C/Mg] and [C/H] are identical. There is a slight slope in [O/Mg] with metallicity, so this could add a systematic to our model.

As we focus on constraining relative yields, we neglect O and Mg yield variations in the main text (excluding the uniform scaling of yields and massloading in Section~\ref{sec:outflows}). There is substantial variation in predicted Mg yields. Most models predict relatively flat trends metallicity (even with rotation as in \citealt{LC18}). However, the variation is significant and our adopted $\Yoc$ yield is much higher than most models. This is a known problem (see \citealt{emily+21}). \cc\ models underpredict $[{\rm Mg}/{\rm O}]$, and the reason why is unknown.

\begin{figure*}
    \centering
    \includegraphics{cc_yields.pdf}
    \caption[High-Mass Star Carbon Yields]{
        C yields from high-mass stars.
        \textbf{Left} The \imf-weighted \cc\ yield of C as a function of metallicity.
        \textbf{Right} The \cc\ [C/Mg] abundance ratio, defined in Eq.~\ref{eq:c_mg_cc}. The black line is the derived C yield from Section \ref{sec:equilibrium},
    $\Ycc = 0.0028 + 0.001 (Z/Z_{\odot})$. Yields are shown for tables from 
    \citet[red triangles]{WW95}, \citet[orange squares and diamonds]{sukhbold+16}, 
    \citet[green stars]{NKT13}, and \citet[blue circles]{LC18}. \citet{sukhbold+16} report yields for different black hole landscapes, while \citet{LC18} provide yields at different rotational velocities.
    In the top panel, the pink line denotes $\Ycagb$ from \cxi{} for comparison. All models include wind yields.
}
    \label{fig:y_cc}
\end{figure*}

\section{The Equilibrium Approximation}\label{sec:equilibrium}

In the presence of metal-poor gas accretion and feedback-driven outflows, galaxies reach an equilibrium abundance in which production of new metals is balanced by losses to new stars and outflows \citep{larson72, dalcanton07, FD08, PS11, lilly13}.
While our galaxy is likely not in perfect equilibrium or described by a single, homogeneous chemical
envirnoment, the equilibrium approximation is nevertheless useful in understanding yields and
metallicity dependence of solar neighborhood stars \citep[e.g.][]{james_dwarf,james+23,WAF17}.

We assume a simple \textit{one-zone} chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.  Newly produced metals are homogeneously and instantaneously mixed, so spatial dependence is neglected.
We define $M_{X}$ to be the mass of element $X$ in the gas-phase, $\dot{M}_\star$ to be the star formation rate (in $\Mo\,{\rm yr}^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ (representing the strength of outflows). A \ssp\ returns a fraction $r$ of their birth mass to the interstellar medium, due to ejected stellar envelopes.\footnotemark{} 
Here, we will use $\alpha$ as a representative $\alpha$-element, such as O or Mg. 
Given the \imf-averaged yield $y_\alpha$, the rate of change in the gas-phase mass of $\alpha$ is a simple sum of sources and sinks,
\begin{equation}
    \dot{M}_\alpha =  y_\alpha\,\dot{M}_\star - \dot{M}_{\alpha,\,\text{remnants}} - \dot{M}_{\alpha,\,\text{outflows}}
\end{equation}
where $y_\alpha\,\dot{M}_{\star}$ describes \cc\ enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_\alpha\,\dot{M}_\star\,(1-r)$.  And, the outflows deplete mass at a rate $Z_\alpha \,\dot{M}_\star\,\eta$. (We assume the composition of outflows is the same as the interstellar medium.) Substituting for $\eta$ and $r$,  
\begin{equation}
    \dot{M}_\alpha= y_\alpha\,\dot{M}_\star - (1 + \eta - r)\,Z_\alpha\,\dot{M}_\star.
\end{equation}
Assuming an exponentially declining \sfh{} $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, the equilibrium abundance is derived analytically by setting $\dot{Z}_\alpha=0$.
\begin{equation}\label{eq:z_eq}
    Z_\alpha^{\rm eq}(R) = \frac{y_\alpha}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}},
\end{equation}
where $\tau_{\star}$ is the star formation rate. ($\eta$ depends on $R$ to create a metallicity gradient. See also Section \ref{sec:vice} and Eq.~\ref{eq:eta_r}.)
In the special case of constant star formation, $\tau_{\rm sfh}\to\infty$, the denominator simplifies to $1+\eta-r$.

Likewise, for \agb\ stars,%
\footnote{In detail, the \agb\ contribution requires integration over the \sfh\ to account for the finite lifetimes of stars. However, for this exponential \sfh, this effect is minimal, so we simply use $\Ycagb$ for the current \agb\ contribution.}
Eq.~\ref{eq:z_eq} then becomes
\begin{equation}
    Z_{\rm C}^{\rm eq}(R) = \frac{\Ycc + \Ycagb}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}}
\end{equation}
And the equilibrium C/$\alpha$ abundance ratio is
\begin{equation}\label{eq:z_co}
    \frac{Z_{\rm C}^{\rm eq}}{Z_\alpha^{\rm eq}} = \frac{\Ycc + \Ycagb }{y_\alpha}.
\end{equation}
Analogous to \cite{james+23} arguments about N, the trends in abundance ratios are set by yield ratios in these \gce{} models. The effect of other \gce{} parameters (most importantly $\eta$) cancels. As a consequence, yield ratios should establish abundance ratio trends in models which assume a different normalization of element yields and mass-loading (see discussion below).
This argument can also be inverted to infer  yields from abundance ratio trends. To the extent that observed C and $\alpha$ trends reflect the equilibrium abundances in different Galactic regions, we can infer the \cc\ yield given an assumed \agb\ star yield (or vice versa). Inferring $\Ycc$ from $\Ycagb$, 
\begin{equation}
    y_\text{C}^\alpha =  y_\alpha^\text{CC} \frac{Z_\text{C,\,eq}}{Z_{\alpha,\,{\rm eq}}} - \Ycagb
\end{equation}
Rewriting this expression as a relative yield of C and Mg,
\begin{equation}\label{eq:y_c_eq}
    \frac{\Ycc}{y_\alpha} = \left(\frac{Z_{\text{C}}}{Z_{\alpha}}\right)_\odot 10^{[{\rm C}/\alpha]} - \frac{\Ycagb}{y_\alpha}.
\end{equation}

\footnotetext{$r\approx0.4$ for a \citealt{kroupa01} \imf.}

\subsection{Analytic Yield Models}

As we will discuss in Section \ref{sec:outflows}, the normalization of yields and $\eta$ is degenerate. This can be observed in Eq.~\ref{eq:z_co}, where changes in $\eta$ or the scaling of $y_{\rm C}/y_{\rm Mg}$ would not affect equilibrium trends. Furthermore, in Eq.~\ref{eq:z_eq}, an increase in both $y_{\rm Mg}$ and $\eta$ would leave $Z_{\rm Mg}^{\rm eq}$ unchanged. Our models here are unable to distinguish the overall scaling of yields and outflow mass loading. Choosing to keep $y_{\rm C}$ fixed reduces unnecessary free parameters. So, at solar metallicity, we set
\begin{subequations}
    \begin{align}
        \Yct(\Zo) &= \Ycc + \Ycagb = 0.005 \\
        \left(\frac{\Yct}{\Yoc}\right)_{\Zo} &= 2.7.
    \end{align}
\end{subequations}
This ratio results in an equilibrium abundance ${\rm [C/Mg]} = -0.09$, which is consistent with the subgiant sample and is within \about{20\%} of the solar C/Mg mixture from \citet{asplund+09}.

In Section \ref{sec:agb_results}, we will show that none of the four \agb\ yield sets (\cxi{}, \kx{}, \vxiii{}, \kxvi{}) produce enough C relative to this $\Ycc$ value. So, we introduce normalization factors, $\alpha_{\rm agb}$, and $\alpha_{\rm CC}$ which denote a multiplicative scaling of $\Ycc$ and $\Ycagb$ respectively. 
\begin{subequations} \label{eq:alpha}
    \begin{align}
        \Ycagb &\rightarrow \alpha_{\rm agb}\ \Ycagb \\
        \Ycc &\rightarrow \alpha_{\rm cc}\ \Ycc\\
        \alpha_{\rm CC} &= \frac{\Yct - \alpha_{\rm agb}\ \Ycagb}{\Ycc}
    \end{align}
\end{subequations}
Above, $\alpha_{\rm cc}$ is dependent on the choice of $\alpha_{\rm agb}$ to keep the total C yield constant.

In Section \ref{sec:agb_results}, we find that $\alpha_{\rm agb}\approx 2.9$ for the \cxi{} yield model. Using this \agb\ yield, we can now use Eq.~\ref{eq:y_c_eq} to find an estimate of $\Ycc$ as a function of metallicity. We show the values of $\Ycc$ we obtain from the subgiant sample in Fig.~\ref{fig:analytic}. From a regression analysis, we suggest here that
\begin{equation}\label{eq:zeta}
    \Ycc = 0.004 \left(1 - \zeta + \zeta\,\frac{2Z}{Z + Z_\odot} + \frac{2\xi}{1 + Z/Z_{\rm III}}\right),
\end{equation}
where $\zeta\approx0.001$ describes the metallicity dependence.

>>>>

\graphicspath{{figures/}}

@@

\newcommand{\Yct}{\ensuremath{y_{\rm C}}}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm cc}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm cc}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm agb}}}
\newcommand{\y}{\ensuremath{\rotatebox[origin=B,y=0.5ex]{180}{y}}}

\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else M$_{\sun}$
    \fi}
\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}

@@

After a single stellar population (SSP)\footnotemark{} forms, \cc{} are the first chemical enrichers. \cc{} explode within $\lesssim 40$\,Myr, providing light elements (e.g. C, O, and Mg) and heavier elements (Fe and beyond). Next, low-mass stars begin to reach the end of their lives, entering the \agb{} phase. By shedding their outer layers, \agb{} stars are important sources of C, N, and neutron capture elements.  Finally, white dwarfs explode in \ia{}, releasing Fe and other iron-peak elements.

@@

To quantify yields, we define the stellar yield to be the fraction of a stars initial mass which is newly synthesized and released as a given element. 
For an element $X$ and star with mass $M$, the net-fractional stellar yield $\y$ is 
\begin{equation}
\y_{X} =  \Delta Z_X \frac{M_{\rm ejected}}{M_{\rm birth}}
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm birth}$  are the total ejected mass and the birth mass of the star, and $\Delta Z_X$ is the change in $Z_X$ from the birth material to the ejected material of the star.%
\footnote{$Z_{X}$ represents the mass fraction of element $X$.}
%
For example, a 1\,\Mo\ star with $\y_{\C} = 10^{-3}$ will add $10^{-3}\,\Mo$ of new C to the interstellar medium. 
Also, note that yields may be negative if the material returned to the interstellar medium has a lower abundance $Z_X$ than the material the star was formed from.
Although per-star yields are necessary to compute \agb{} star enrichment rates in \gce{}  models, \imf-averaged yields are useful in interpreting their predictions. 
Given the fractional yield $\y(M, Z)$ as a function of initial stellar mass $M$ and metallicity $Z$, the \imf-averaged yield is given by 
\begin{equation} \label{eq:imf-yield}
    y_{\rm X}(Z,t) = 
    \frac{
    \int_{M_{\rm min}(t)}^{M_{\rm max}} 
    \y_{\rm X}(M, Z)
    \frac{dN}{dM}\ M\ dM
}
{
    \int \frac{dN}{dM}\ M\ dM
}
\end{equation}
where ${dN}/{dM}$ is the \imf, $M_{\rm max}=100\,\Mo$ is the maximum stellar mass, and $M_{\rm min}(t)$ is the mass of stars with lifetime $t$.\footnotemark{}
We use $t=10\,$Gyr for total yields when $t$ is not used.
To calculate the \imf-averaged net yields, we use the Versatile Integrator for Chemical Evolution code (\VICE).\footnotemark{}

@@

	\begin{tabular}{l l l l}
		\hline
        Element & $y^{\rm cc}$ & $\y^{\rm agb}$ & $y^{\rm ia}$ \\
		\hline
        C & Eq.~\ref{eq:zeta} & $2.9\times{\rm C11}$ &  0 \\
        O & 0.015 & 0 & 0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than \about{2}\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
Between \about{2} and \about{5} \Mo, third dredge up becomes important, enriching the outer layers with C, making these stars the most important producers of C. 
In \agb\ stars more massive than \about{5}\,\Mo, both hot bottom burning and third dredge up occur; however, third dredge up is much more efficient, resulting in significant \C[12] destruction.

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following,
\begin{description}
    \item \cxi: \citet{cristallo+11, cristallo+15}
    \item \kx: \citet{karakas10}
    \item \vxiii: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \kxvi: \citet{KL16, karakas+18}
\end{description}
In Appendix~\ref{sec:oob_models}, the masses, metallicites, and notable differences between each yield set are briefly described. 
For our models to match observations, we find that need to uniformly amplify these yield tables. We use \cxi\ table, amplified by a factor of 2.9, as the fiducial \agb\ yield.

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yield for these four models.
Most models agree on the qualitative shape of the net fractional \agb\ C yield.
Stellar yields peak between masses of about 2--4 \Mo and decline as stars become more or less massive. As metallicity increases, the total \agb\ C yield decreases. The mass of peak C yields also increases slightly with metallicity. More metal-rich stars lose mass more quickly due to higher opacities, so these stars behave more like low-mass AGB stars. MORE PHYSICS

@@

The right panel of Fig.~\ref{fig:agb-ssp} shows \imf-averaged C yields for each \agb\ model as a function of metallicity.
\vxiii{} differs in that it shows a non-monotonic metallicity dependence. However, this effect is only for models with $\log Z/Z_\odot \lesssim -1$.
Otherwise, models differ only in their yield normalization and metallicity dependence. 
All yield models span a range of \about{2} for a given metallicity.
For example, the three models \cxi, \kx{}, and \kxvi{} predict $y_\text{C}^\text{agb}$ to be between 0.006 and 0.008 at solar metallicity, but \cxi\ has a much shallower metallicity dependence than the \kx{} and \kxvi{} models. \vxiii{} instead predicts a yield \about{0.004}.

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process. However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the complexity of stellar modeling.

Fig.~\ref{fig:y_cc} plots calculations of the \imf-integrated yields, defined with Eq.~\ref{eq:imf-yield} (computed using \VICE's \texttt{vice.yields.ccsne.fractional} function). 
\cc{} models predict a wide range of C yields, spanning almost a factor of ten. 
Both the \citet{NKT13} and \cite{LC18} models show positive metallicity dependence. 
As metallicity increases, stars lose more of their mass to winds. In particular, C enriched envelop material is lost through winds before synthesized into heavier elements, so C yields can be strongly metallicity dependent (VERIFY).
Fig.~\ref{fig:y_cc} shows the \cxi{} \agb\ model for comparison on the left. Especially at $Z\approx Z_\odot$, most \cc{} models dominate \agb\ C production. Later, we will also show empirically this is the case.
The right of Fig.~\ref{fig:y_cc} shows the \cc{} [C/Mg] ratio for the different models, defined by
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/Mg]^{CC}} = \log_{10}\left( \frac{\Ycc}{\Yoc}\right) - \log_{10} \left( \frac{Z_{{\rm C},\ \sun }}{Z_{{\rm Mg},\ \sun }} \right).
\end{equation}
If only \cc{} produced C, then ${\rm [C/Mg]^{CC}}$ describes the equilibrium abundance of [C/Mg].
Different \cc\ models also span a large range in [C/Mg]. 
We chose to instead parameterize $\Ycc$ to simplify model inputs, as most \cc\ models fail to achieve near-solar [C/Mg].

Both rotation and explodability introduce substantial variations. The \cite{LC18} models include rotation, showing that variations in the rotational velocity of the star can dramatically increase the magnitude and metallicity dependence of $\Ycc$. Rotation induces more mixing allowing the CO core to grow larger and contributes to wind losses. As we will later show, \cc\ C production needs to be strongly metallicity-dependent at $Z/Z_\odot \approx 1$, which is consistent with the \cite{LC18} rapidly rotating models.
Assumptions about the explodability landscape affect C and Mg production. Increasing the fraction of stars that explode increases $\Ycc$, as stars that directly collapse do not contribute to explosive yields \citep{emily+21}. However, C is relatively unaffected by the black-hole landscape, as very massive stars contribute C through enriched winds. Since Mg is formed deeper in the core of \hms, the Mg yield drops much more steeply, so models where few stars explode (S16/W18) have higher [C/Mg].

O and Mg are both $\alpha$-elements, light elements produced in \cc{} through He-nuclei fusion. 
As we focus on constraining relative yields, we neglect O and Mg yield variations in the main text (excluding the uniform scaling of yields and massloading in Section~\ref{sec:outflows}). There is substantial variation in predicted Mg yields. 
Most models predict relatively flat trends metallicity (even with rotation as in \citealt{LC18}). 
However, the variation is significant and our adopted $\Yoc$ yield is much higher than most models, but O and Mg yields of \cc\ models do not fully match observations. \cc\ models underpredict $[{\rm Mg}/{\rm O}]$, and the reason why is unknown (see e.g. \citealt{emily+21}). Here, we assume ${\rm [O/Mg]} = 0$, which is not compatible with \cc\ models but is consistent with \apogee{} observations \citep{weinberg+19, weinberg+22}.
Because O and Mg are not perfectly identical in \apogee{}, the specific choice of $\alpha$-element may be a possible systematic.

\begin{figure*}
    \centering
    \includegraphics{cc_yields.pdf}
    \caption[High-Mass Star Carbon Yields]{
        C yields from high-mass stars.
        \textbf{Left} The \imf-weighted \cc\ yield of C as a function of metallicity.
        \textbf{Right} The \cc\ [C/Mg] abundance ratio, defined in Eq.~\ref{eq:c_mg_cc}. The black line is the derived C yield from Section \ref{sec:equilibrium},
    $\Ycc = 0.0028 + 0.001 (Z/Z_{\odot})$. Yields are shown for tables from 
    \citet[red triangles]{WW95}, \citet[orange squares and diamonds]{sukhbold+16}, 
    \citet[green stars]{NKT13}, and \citet[blue circles]{LC18}. \citet{sukhbold+16} report yields for different black hole landscapes, while \citet{LC18} provide yields at different rotational velocities.
    In the top panel, the pink line denotes $\Ycagb$ from \cxi{} for comparison. All models include wind yields. 
    (OUT OF DATE)
}
    \label{fig:y_cc}
\end{figure*}

\section{Analytic Approximations}
In this section, we describe approximations of galactic abundance trends and C yield models. While these are insufficient to describe the galaxy in full detail, we can use these simplified models to inform our investigation.

\subsection{Equilibrium Abundances}\label{sec:equilibrium}

Galaxies, when moderated by metal-poor gas accretion and feedback-driven outflows, reach a chemical equilibrium. The production of new metals is balanced by losses to new stars and outflows \citep{larson72, dalcanton07, FD08, PS11, lilly13}.
While our galaxy is likely not in perfect equilibrium or described by a single, homogeneous chemical
envirnoment, the equilibrium approximation is nevertheless useful in understanding yields and
metallicity dependence of solar neighborhood stars \citep[e.g.][]{james_dwarf,james+23,WAF17}.

Here, we assume a simple \textit{one-zone} chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.  Newly produced metals are homogeneously and instantaneously mixed, so this model evolves as a single, spatially-uniform chemical zone.
We define $M_{X}$ to be the mass of element $X$ in the gas-phase, $\dot{M}_\star$ to be the star formation rate (in $\Mo\,{\rm yr}^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ (representing the strength of outflows). A \ssp\ returns a fraction $r$ of their birth mass to the interstellar medium, due to ejected stellar envelopes.%
\footnote{$r\approx0.4$ for a \citealt{kroupa01} \imf.}
Here, we will use $\alpha$ as a representative $\alpha$-element, such as O or Mg. 
Given the \imf-averaged yield $y_\alpha$, the rate of change in the gas-phase mass of $\alpha$ is a simple sum of sources and sinks,
\begin{equation} \label{eq:mdot_1}
    \dot{M}_\alpha =  y_\alpha\,\dot{M}_\star - \dot{M}_{\alpha,\,\text{remnants}} - \dot{M}_{\alpha,\,\text{outflows}}
\end{equation}
where $y_\alpha\,\dot{M}_{\star}$ describes \cc\ enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_\alpha\,\dot{M}_\star\,(1-r)$.  And, the outflows deplete mass at a rate $Z_\alpha \,\dot{M}_\star\,\eta$. (We assume the composition of outflows is the same as the interstellar medium.) With these definitions, Eq.~\ref{eq:mdot_1} is
\begin{equation}
    \dot{M}_\alpha= y_\alpha\,\dot{M}_\star - (1 + \eta - r)\,Z_\alpha\,\dot{M}_\star.
\end{equation}
Assuming an exponentially declining \sfh{} $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, the equilibrium abundance is derived analytically by setting $\dot{Z}_\alpha=0$.
\begin{equation}\label{eq:z_eq}
    Z_\alpha^{\rm eq}(R) = \frac{y_\alpha}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}},
\end{equation}
where $\tau_{\star}$ is the star formation rate. ($\eta$ depends on $R$ to create a metallicity gradient. See also Section \ref{sec:vice} and Eq.~\ref{eq:eta_r}.)
In the special case of constant star formation, $\tau_{\rm sfh}\to\infty$, the denominator simplifies to $1+\eta-r$.

Likewise, for C production, where both \agb\ stars and \cc\ play a role,%
\footnote{In detail, the \agb\ contribution requires integration over the \sfh\ to account for the finite lifetimes of stars. However, for this exponential \sfh, this effect is minimal, so we simply use $\Ycagb$ for the current \agb\ contribution.}
Eq.~\ref{eq:z_eq} then becomes
\begin{equation}
    Z_{\rm C}^{\rm eq}(R) = \frac{\Ycc + \Ycagb}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}}
\end{equation}
The equilibrium C/$\alpha$ abundance ratio is then
\begin{equation}\label{eq:z_co}
    \frac{Z_{\rm C}^{\rm eq}}{Z_\alpha^{\rm eq}} = \frac{\Ycc + \Ycagb }{y_\alpha}.
\end{equation}
Analogous to \cite{james+23} arguments about N, the trends in abundance ratios are set by yield ratios in these \gce{} models. The effect of other \gce{} parameters (most importantly $\eta$) cancels. As a consequence, yield ratios should establish abundance ratio trends in models which assume a different normalization of element yields and mass-loading (see discussion below).
This argument can also be inverted to infer  yields from abundance ratio trends. To the extent that observed C and $\alpha$ trends reflect the equilibrium abundances in different Galactic regions, we can infer the total C yield.
\begin{equation}\label{eq:y_c_eq}
    \Ycc + \Ycagb =  y_\alpha \left(\frac{Z_{\rm C}}{Z_{\alpha}}\right)_\odot 10^{[{\rm C}/\alpha]}.
\end{equation}
As we know [C/$\alpha$] as a function of metallicity, Eq.~\ref{eq:y_c_eq} enables us to estimate both the magnitude and metallicity-dependence of $\Yct/y_\alpha$.
In Fig.~\ref{fig:analytic}, we show the inferred total C yields, based on this equation, and our best fitting linear model. From a linear regression, we suggest that
\begin{subequations}\label{eq:yc_inferred}
    \begin{align}
        \frac{\Yct(Z)}{y_{\rm O}} &\approx \frac{1}{3} + 4 \left(Z-\Zo\right) \\
        \frac{\Yct(Z)}{y_{\rm Mg}} &\approx 2.7 + 32 \left(Z-\Zo\right).
    \end{align}
\end{subequations}
These yield ratios results in an equilibrium abundance $[{\rm C}/\alpha] = -0.09$, which is consistent with the subgiant sample and is within \about{20\%} of the solar C/Mg mixture from \citet{asplund+09}.

\subsection{Analytic AGB}

In Section \ref{sec:agb_results}, we will show that none of the four \agb\ yield sets (\cxi{}, \kx{}, \vxiii{}, \kxvi{}) produce enough C relative to this $\Ycc$ value. So, we introduce a normalization factor $\alpha_{\rm agb}$, which denotes a multiplicative scaling of $\Ycagb$ 
\begin{equation} \label{eq:alpha}
        \Ycagb \rightarrow \alpha_{\rm agb}\ \Ycagb.
\end{equation}
To parameterize the \agb\ contribution to C production, we define $f_{\rm agb}$ to be the fraction of C which comes from \agb\ stars. 
\begin{equation}\label{eq:f_agb}
    f_{\rm agb} \equiv \frac{\Ycagb(Z=\Zo)}{\Yct(Z=\Zo)},
\end{equation}
($f_{\rm agb}$ accounts for the multiplicative factor $\alpha_{\rm agb}$ as defined in Eq.~\ref{eq:alpha}).
